---
title: 'COVID-19 Rare disease Phenotype: characteristics of individuals'
author: 'Chris Tomlinson, Johan Thygesen'
date: 'Knitted on `r date()`'

output:
  html_document:
    df_print: paged
---

**Description** 

This notebook aggregates the data within `ccu019_paper_cohort_rare_disease` and stratifies by `severity` which represents a picture of the most severe healthcare presentation per patient (mutually-exclusive).
  
**Project(s)** CCU019

**Author(s)** Chris Tomlinson, adapted to CCU019 by Johan Thygesen
 
**Reviewer(s)** 
 
**Date last updated** 2023-01-19
 
**Date last reviewed** *NA*
 
**Date last run** `r date()`
 
**Data input**  
* `ccu013_paper_table_one_56million_denominator`

**Data output**  
* Self-contained as the knitted `.html` file from this notebook

**Software and versions** `SQL`, `R`
 
**Packages and versions**

```{r config}
# Load libraries prior to sessionInfo() to ensure documented
library(dplyr)
library(tidyr)
library(magrittr)
library(stringr)
library(tableone)
library(DBI)

setwd("/mnt/efs/dars_nic_391419_j3w9t_collab/CCU019")

#warning("Requires authenticated connection to Spark saved as 'con'")
source("/mnt/efs/j.thygesen/dbConnect.R")
```

```{r sessionInfo}
sessionInfo()
```


## 1. Load & process cohort

```{r load}
# Load cohort
if(!exists("cohort_small")){load("/mnt/efs/dars_nic_391419_j3w9t_collab/CCU019/data/ccu019_pop_cohort_noID.Rdata")}

```


```{r process}
cohort_small %<>% 
  mutate(sex = case_when(sex == '1' ~ 'Male', 
                         sex == '2' ~ 'Female',
                         sex == '9' ~ 'Unknown',
                         sex == '0' ~ 'Unknown')
         ) %>% 
  # Age: currently no NAs (encoded as 999) or Unknowns, else
  # mutate(age = na_if(age, 999)) %>% 
  mutate(age = as.numeric(age)) %>% 
  # Add binned ages, as per descriptive paper
  mutate(age_bin = case_when(age < 18 ~ "Under 18",
                           age <= 29 ~ "18-29",
                           age <= 49 ~ "30-49",
                           age <= 69 ~ "50-69",
                           age >= 70 ~ "70 or over")
       ) %>%  
  select(-age) %>% 
  # Factorise!
  mutate(across(c(sex, 
                  age_bin, 
                  ethnic_group, 
                  IMD_quintile, 
                  high_risk),
                as.factor)
         ) %>% 
  # re-level
  mutate(ethnic_group = relevel(ethnic_group, ref='White')
         ) %>% 
  # Factorise outcomes to get counts & % instead of mean(sd)
  mutate(across(c(severity,
                  death_covid,
                  death_all),
                as.factor)
         ) %>% 
  # Re-order to desired structure for tables
  relocate(death_covid,
           death_all,
           sex,
           age_bin,
           ethnic_group,
           IMD_quintile,
           high_risk )
```

## All 58 million: COVID & No COVID
```{r population}
cohort_small %>%
  select(-severity, -rare_disease) %>% 
  CreateTableOne(data=.) %>% 
  print(., 
        format.args = list(big.mark = ",")
        ) %>% 
  kableone(., 
           format.args = list(big.mark = ",")
           )
```

## All with rare disease
```{r rare_disease}
cohort_small %>%
  filter(rare_disease == 1) %>%
  select(-severity, -rare_disease) %>% 
  CreateTableOne(data=.) %>% 
  print(., 
        format.args = list(big.mark = ",")
        ) %>% 
  kableone(., 
           format.args = list(big.mark = ",")
           )
```

## Stratified rare_disease by severity
```{r rare_disease_severity}
cohort_small %>%
  filter(rare_disease == 1) %>%
  filter(severity %in% c("2_hospitalised", "3_ventilatory_support"))  %>%
  select(-rare_disease) %>% 
  CreateTableOne(data=.,
                 strata = "severity") %>% 
  print(., 
        format.args = list(big.mark = ",")
        ) %>% 
  kableone(., 
           format.args = list(big.mark = ",")
           )
```

## COVID ONLY
```{r covid_only}
cohort_small %>%
  filter(rare_disease == 1) %>%
  filter(severity != 'no_covid') %>% 
  select(-severity, -rare_disease) %>% 
  CreateTableOne(data=.) %>% 
  print(., 
        format.args = list(big.mark = ",")
        ) %>% 
  kableone(., 
           format.args = list(big.mark = ",")
           )
```

## ALL COVID Deaths ONLY
As stratifying by severity only give those individuals where deaths was their only healthcare presentation, we want to show the demographics of all the COVID-19 deaths too.
```{r all_covid_deaths}
cohort_small %>%
  filter(death_covid == 1) %>% 
  filter(rare_disease == 1) %>%
  select(-severity) %>% 
  CreateTableOne(data=.) %>% 
  print(., 
        format.args = list(big.mark = ",")
        ) %>% 
  kableone(., 
           format.args = list(big.mark = ",")
           )
```